### 2022-09-29 main/chapter_theory.tex
write a proper preface

Early versions of the Standard Model had only two fermion generations and
no neutrino masses~\cite{
wells2020discovery,
bjorken1985november
},
but the third generation and non-zero neutrino masses are now observed;
for example, the third generation is shown by charge-parity $CP$ asymmetry in
kaons~\cite{
cronin1964evidence,
kabayasji1973cpv
},
precise measurements of the $Z$ boson~\cite{
lep2006precision
},
and direct observations~\cite{
perl1977evidence,
herb1977observation,
abachi1995observation
};
neutrino masses are also necessary to explain their observed flavour
oscillations~\cite{
kamiokande1998measurement,
superk1998evidence,
superk1999measurement,
lsnd1998evidence
}.

@@

Other mysteries of cosmology include:
the small positive cosmological constant,
the inflation of the early universe,
and baryogenesis~\cite{
wells2020discovery,
riess1998observational
}.
Solutions to these mysteries all lie somewhere between uncertain and unknown,
but may not be entirely out of reach;
for example, the Higgs field could plausibly have driven
inflation~\cite{bezrukov2008higgs},
and the Standard Model does slightly violate $C$ and $CP$, both of which
must be broken for baryogenesis~\cite{sakharov1991re}.

>>>>

Early versions of the Standard Model had only two fermion generations and
no neutrino masses~\cite{
wells2020discovery,
bjorken1985november
},
but the third generation and non-zero neutrino masses are now observed;
for example, the third generation is shown by charge-parity ($\mathrm{CP}$
asymmetry in kaons~\cite{
cronin1964evidence,
kabayasji1973cpv
},
precise measurements of the $Z$ boson~\cite{
lep2006precision
},
and direct observations~\cite{
perl1977evidence,
herb1977observation,
abachi1995observation
};
neutrino masses are also necessary to explain their observed flavour
oscillations~\cite{
kamiokande1998measurement,
superk1998evidence,
superk1999measurement,
lsnd1998evidence
}.

@@

Other mysteries of cosmology include:
the small positive cosmological constant,
the inflation of the early universe,
and baryogenesis~\cite{
wells2020discovery,
riess1998observational
}.
Solutions to these mysteries all lie somewhere between uncertain and unknown,
but may not be entirely out of reach;
for example, the Higgs field could plausibly have driven
inflation~\cite{bezrukov2008higgs},
and the Standard Model does slightly violate $\mathrm{C}$ and $\mathrm{CP}$,
both of which must be broken for baryogenesis~\cite{sakharov1991re}.

### 2022-09-29 main/frontmatter.tex
write a proper preface

Data are best interpreted with an understanding of their origins, so we begin
by introducing the theoretical and experimental background to our search, and
the subtleties of our data-analysis procedures in their historical context.
\end{abstract}

\begin{singlespacing}
\begin{declaration}
\noindent
This thesis is the result of my own work and includes nothing which is the
outcome of work done in collaboration except as
%declared in the preface and
specified in the text.
%
It is not substantially the same as any work that has already been submitted
before for any degree or other qualification.
%except as declared in the preface and specified in the text.
%
It includes neither more nor fewer than
two squared times three cubed
%one hundred and eight
jelly beans.
%
It does not exceed the prescribed word limit for the
Physics and Chemistry
Degree Committee.
\vspace*{1cm}
\begin{flushright}
Rupert~Tombs
\end{flushright}
\end{declaration}
\end{singlespacing}

@@

\paragraph{\atlas.}
Thank you to the Cambridge \atlas\ group and especially Tina~Potter for
sharing help and wisdom.
Thanks to all collaborators mentioned in Chapter~\ref{chapter:2ljets},
to Ben~Hooberman and Sarah~Williams for leadership, and especially to
Jon~Long for our teamwork in the push for publication,

\paragraph{Misc.}
Thank you Christopher~Gorham~Lester, Holly~Pacey, and Daniel~No\"el for your
stunningly precise and valuable comments on drafts of this thesis.
Thank you to Homerton College and Melanie~Keene.
This document uses parts of the `hepthesis' class~\cite{hepthesis}.
\end{acknowledgements}

\begin{preface}
\begin{singlespacing}
\begin{epigraphs}
\qitem{%
Begin reading at Chapter N. Do \emph{not} read the quotations
that appear at the beginning of the chapter.%
}%
{Donald~E.~Knuth,
\textit{The Art of Computer Programming, Volume~1},
1975~\cite{knuth1975art}}
\end{epigraphs}
\end{singlespacing}

Data Intensive Science in High-Energy Physics

\TODO{Context of Standard Model theory, directions in which future developments
could improve it, and supersymmetric models}
supersymmery (SUSY)
Chapter~\ref{chapter:theory}

\TODO{the design of the \atlas\ detector}
Chapter~\ref{chapter:experiment}

\TODO{theory and practice of data science, as it could be for clarity
and how it is in our \atlas\ SUSY conventions}
Chapter~\ref{chapter:searches}

\TODO{the main event: the search with many of its gory details in addition
to those presented publically in the paper~\cite{atlas2022searches}}
Chapter~\ref{chapter:2ljets}

\paragraph{Other work.}
\TODO{Use some of the data analysis ideas of Chapter~\ref{chapter:searches}
to build predictive models
and test them to test for symmetries}
\TODO{citations}

\end{preface}

>>>>

Data are best interpreted with an understanding of their origins, so we begin
by describing the theoretical and experimental background to our search.
Our results depend strongly on the interpretation of data by conventional
analysis procedures, whose properties and explanations are not obvious.
To explain these procedures, we therefore also describe their historical
context and the basic theory of data analysis.
\end{abstract}

\begin{singlespacing}
\begin{declaration}
\noindent
This thesis is the result of my own work and includes nothing which is the
outcome of work done in collaboration except as
%declared in the preface and
specified in the text.
%
It is not substantially the same as any work that has already been submitted
before for any degree or other qualification.
%except as declared in the preface and specified in the text.
%
It includes neither more nor fewer than
two squared times three cubed
%one hundred and eight
jelly beans.
%
It does not exceed the prescribed word limit for the
%
Physics and Chemistry
%
Degree Committee.
\vspace*{1cm}
\begin{flushright}
Rupert~Tombs
\end{flushright}
\end{declaration}
\end{singlespacing}

@@

\paragraph{\atlas.}
Thank you to the whole Cambridge \atlas\ group, and especially Tina~Potter for
sharing help and wisdom.
Thanks to all collaborators mentioned in Chapter~\ref{chapter:2ljets},
to Ben~Hooberman and Sarah~Williams for leadership, and especially to
Jon~Long for our teamwork in the push for publication,

\paragraph{Misc.}
Thank you Christopher~Gorham~Lester, Holly~Pacey, and Daniel~No\"el for your
stunningly precise and valuable comments on drafts of this thesis.
Thank you to Homerton College and Melanie~Keene.
This work was supported by the Science and Technology Facilities Council.
The template for this document derives from `hepthesis'~\cite{hepthesis}.
\end{acknowledgements}

\begin{preface}
\begin{singlespacing}
\begin{epigraphs}
\qitem{%
Begin reading at Chapter N. Do \emph{not} read the quotations
that appear at the beginning of the chapter.%
}%
{Donald~E.~Knuth,
\textit{The Art of Computer Programming, Volume~1},
1975~\cite{knuth1975art}}
\end{epigraphs}
\end{singlespacing}
\noindent
``Two leptons, jets, and missing transverse momentum'' of our title
refers an experimental signature of new phenomena which was targetted by our
recent search in \atlas\ data~\cite{atlas2022searches}.
The subtitle ``yet~more non-evidence for supersymmetry'' refers to how its
negative result is consistent with all other research in the area.
The wording deliberately avoids suggesting evidence against supersymmetry
(SUSY);
we see evidence against specific supersymmetric models, but the abstract
concept of supersymmetry remains at large.

Chapter~\ref{chapter:2ljets} dissects this search project and my contributions
to it.
Earlier chapters introduce relevant content:
Chapter~\ref{chapter:searches} details data analysis,
its theory and conventional practice,
Chapter~\ref{chapter:experiment} describes the design of the \atlas\ detector,
and
Chapter~\ref{chapter:theory} introduces the theory of Standard Model, its
promising future directions, and supersymmetry.
Please consider reading them in forwards order.

Another product of this PhD (not further explored in this thesis),
was the development of a general method for testing symmetries in data~\cite{
lester2021stressed,
tombs2021which
}
and its application to parity violation in hadron colliders~\cite{
lester2022hunting,
Lester:2019bso,
lesterChiralMeasurements2021,
Gripaios:2020hya,
lester2021lorentz
}.
The result is a practical solution which employs the principles of predicting
data and reporting likelihoods as advocated by Chapter~\ref{chapter:searches}.
It uses `self-supervised' machine learning to train predictive models on a
task that can only be solved if the proposed symmetry is violated~\cite{
tombs2021which,
pmlr-v139-radford21a,
Noroozi2016jigsaw,
multitaskself2017,
devlin2019bert
}.
\end{preface}
